---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r load_packages, message = FALSE}
library(dplyr)
library(lattice)
#loading the raw data
myRaw <- read.csv("activity.csv")
```
### NA's must be omitted as requested. keeps rawdata.
```{r}
file <- na.omit(myRaw)
```

## TASK 1 - What is mean total number of steps taken per day?
```{r hist}
#Vector of steps per day
DailyStepsVec <- rowsum(file$steps,file$date)
hist(DailyStepsVec, xlab="Number of steps", ylab="Count", main="Total number of steps taken each day", col = 5)
#will use sprintf to round those values below
MyMean <- mean(DailyStepsVec)
MyMedian <- median(DailyStepsVec)
```
##### The daily mean of the steps is `r sprintf("%.2f", MyMean)` and the median is `r sprintf("%.2f", MyMedian)`

## TASK 2 - What is the average daily activity pattern?

```{r plot}
myInterval <- tapply(file$steps, file$interval, mean)
plot(row.names(myInterval), myInterval, type = "l", xlab = "Interval", 
    ylab = "Average across all Days", main = "Average number of steps", 
    col = "blue")
```

#### The highest average step count is in interval `r names(which.max(myInterval))`

## TASK 3 - Imputing missing values
```{r missing}
missing <- sum(is.na(myRaw))
```

#### There are `r missing` missing values in the raw file

```{r imputing}
#will complete NA values with the 5 min avg
avgSteps <- aggregate(steps ~ interval, data = myRaw, FUN = mean)
#NAs vector
newNAVec <- numeric()
for (i in 1:nrow(myRaw)) {
    myObs <- myRaw[i, ]
    if (is.na(myObs$steps)) {
        steps <- subset(avgSteps, interval == myObs$interval)$steps
    } else {
        steps <- myObs$steps
    }
    newNAVec <- c(newNAVec, steps)
}
#Create a new dataset that is equal to the original dataset but with the missing data filled in
newData <- myRaw
newData$steps <- newNAVec
#make a new hist with the new data with the median and mean
#Vector of steps per day
DailyStepsVec2 <- rowsum(newData$steps,newData$date)
hist(DailyStepsVec2, xlab="Number of steps", ylab="Count", main="Total number of steps taken each day", col = 3)
#will use sprintf to round those values below
MyMean2 <- mean(DailyStepsVec2)
MyMedian2 <- median(DailyStepsVec2)
strMean <- if (MyMean == MyMean2) "the same of " else "different then "
strMedian <- if (MyMedian == MyMedian2) "the same of " else "different then "

```
##### The new daily mean of the steps is `r sprintf("%.2f", MyMean2)` and it's `r sprintf("%s",strMean)` the previous value (`r sprintf("%.2f", MyMean)`)
##### The new daily median of the steps is `r sprintf("%.2f", MyMedian2)` and it's `r sprintf("%s",strMedian)` the previous value (`r sprintf("%.2f", MyMedian)`)

## TASK 4 - Are there differences in activity patterns between weekdays and weekends?
```{r weekdays}
myRaw$date <- as.Date(myRaw$date)
day <- weekdays(myRaw$date)
daylevel <- vector()
for (i in 1:nrow(myRaw)) {
    daylevel[i] <- (if (day[i] == "Saturday" || day[i] == "Sunday" ) "Weekend" else "Weekday")
}
myRaw$daylevel <- daylevel
myRaw$daylevel <- factor(myRaw$daylevel)

StepsAgg <- aggregate(steps ~ interval + daylevel, data = myRaw, mean)
names(StepsAgg) <- c("interval", "daylevel", "steps")
```
### Comparing patterns
```{r xyplot}
xyplot(steps ~ interval | daylevel, StepsAgg, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```
